---
title: "Extreme Dependence Data Analysis"
author: "Alex Reyes"
date: "2024-10-08"
output: html_document
---

***Cleaning the Workspace/Loading Required libraries***
```{r}
# Clean the workspace
rm(list = ls())

getwd()

# Load required packages
library(tidyverse)
library(tidyselect)
library(tidyquant)
library(quantmod)
library(timetk)
library(copula)
library(VineCopula)
library(cowplot)
library(infotheo)
library(GGally)
library(nortest)
library(car)
library(knitr)
library(kableExtra)
library(pander)
library(patchwork)
library(rugarch)
library(FinTS) 
library(reshape2)
library(viridis)
```

***Functions for Stock Closing Prices/Log Returns/Wide-Format for Copula Analysis/plotting prices and log returns***
```{r}
# Create functions for tick marks
number_ticks <- function(n) { 
  function(limits) pretty(limits, n) 
}

date_ticks <- function(n) {
  function(limits) {
    pretty(seq(limits[1], limits[2], length.out = n), n)
  }
}

# Function to get and process stock prices
get_prices <- function(tickers, start_date, end_date, sector_name) {
  prices <- tq_get(tickers, from = start_date, to = end_date) |> 
    mutate(
      symbol = factor(symbol),
      sector = sector_name
    ) |> 
    group_by(symbol) |> 
    mutate(
      close_scaled = as.vector(scale(close)),  # Scale closing prices
      close_lgret = log(close / lag(close))  # Calculate log returns
    ) |> 
    ungroup() |> 
    drop_na() |> 
    select(date, sector, symbol, close, close_scaled, close_lgret)
  
  return(prices)
}

# Function to convert prices data to wide format for copula analysis
get_wide_prices <- function(prices) {
  wide_prices <- prices |> 
    select(date, symbol, close_lgret) |> 
    pivot_wider(names_from = symbol, values_from = close_lgret) |> 
    select(-date)
  
  return(wide_prices)
}

# Define start and end dates for the data
start_date <- "2016-01-01"
end_date <- "2023-12-31"

# Define tickers for the consumer discretionary sector
staples_tickers <- c(
  "COST", "WMT", "KO",
  "PEP", "TGT", "KR",
  "TSN", "CLX", "STZ",
  "MNST", "SYY", "PG")

# Define tickers for the communication services sector
communication_tickers <- c(
  "META", "GOOGL", "NFLX", 
  "DIS", "WBD", "PARA", 
  "CMCSA", "EA", "T", 
  "VZ", "TMUS", "OMC")

# Define tickers for the consumer staples sector
discretionary_tickers <- c(
  "AMZN", "TSLA", "HD",
  "MCD", "LOW", "NKE",
  "SBUX", "MAR", "EBAY",
  "BBY", "ETSY", "MGM")


## Filtering and Processing stocks for Consumer Staples Sector ##


# Fetch and process consumer staples sector stock prices
staples_prices <- get_prices(
  tickers = staples_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Consumer Staples"
)

staples_stable <- staples_prices |> 
  dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
staples_stable$period <- "Stable Market"
staples_stable <- staples_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

staples_pandemic <- staples_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
staples_pandemic$period <- "Pandemic and Downturn"
staples_pandemic <- staples_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods
staples_period <- rbind(
  staples_stable, staples_pandemic
) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Stable Market", "Pandemic and Downturn")
    )
  )


## Filtering and Processing stocks for Consumer Discretionary Sector ##

discretionary_prices <- get_prices(
  tickers = discretionary_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Consumer Discretionary"
)

discretionary_stable <- discretionary_prices |> 
  dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
discretionary_stable$period <- "Stable Market"
discretionary_stable <- discretionary_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

discretionary_pandemic <- discretionary_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
discretionary_pandemic$period <- "Pandemic and Downturn"
discretionary_pandemic <- discretionary_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods ()
discretionary_period <- rbind(
  discretionary_stable, discretionary_pandemic
) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Stable Market", "Pandemic and Downturn")
    )
  )


# Filtering and Processing stocks for Communication Services Sector

communication_prices <- get_prices(
  tickers = communication_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Communication Services"
)

communication_stable <- communication_prices |> 
  dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
communication_stable$period <- "Stable Market"
communication_stable <- communication_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

communication_pandemic <- communication_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
communication_pandemic$period <- "Pandemic and Downturn"
communication_pandemic <- communication_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods ()
communication_period <- rbind(
  communication_stable, communication_pandemic
) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Stable Market", "Pandemic and Downturn")
    )
  )


```


***Testing ARCH Effects, and Fitting one when necessary (significant P-value)***
```{r}


#' Test for ARCH effects and fit a GARCH(1,1) model if necessary
#'
#' This function takes a time series of log-returns (numerical vector) as input, 
#' performs an ARCH test using the FinTS::ArchTest function, and checks whether 
#' a GARCH structure is necessary. If the ARCH test indicates the need for a GARCH model, 
#' the function fits a GARCH(1,1) model using the rugarch package and returns the residuals.
#' If the ARCH test fails (i.e., no GARCH structure is necessary), it simply returns the input time series.
#'
#' @param x A numerical vector containing the time series of log-returns.
#' @param lags The number of lags for the ARCH test (default is 5).
#' @param alpha The significance level for the ARCH test (default is 0.05).
#' 
#' @return If no ARCH effects are found, the function returns the original time series. 
#' If ARCH effects are detected, it returns the residuals of the fitted GARCH(1,1) model.
#'
#' @examples
#' # Example with simulated GARCH(1,1) data:
#' n <- 1000
#' x <- as.numeric(fitted(ugarchpath(ugarchspec(), n.sim = n)))
#' result <- test_and_fit_garch(x)
#' print(result)
#'
#' @import rugarch
#' @import FinTS
#' @export
test_and_fit_garch <- function(x, lags = 5, alpha = 0.05) {
  
  # Step 1: Perform the ARCH test using FinTS::ArchTest
  arch_test <- FinTS::ArchTest(x, lags = lags)
  
  # Extract the p-value from the test
  p_value <- arch_test$p.value
  
  # Step 2: Decision rule based on p-value
  if (p_value >= alpha) {
    # If the p-value is greater than alpha, no ARCH effects are detected
    print("No significant ARCH effects. A GARCH model might not be necessary.")
    return(x)  # Return the original time series
  } else {
    # If the p-value is less than alpha, ARCH effects are detected
    print("There is evidence of ARCH effects. A GARCH model may be necessary.")
    
    # Step 3: Fit the GARCH(1,1) model to the data 'x'
    garch_spec <- ugarchspec(
      variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
      mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
      distribution.model = "norm"
    )
    
    # Fit the model
    garch_fit <- ugarchfit(spec = garch_spec, data = x)
    
    # Step 4: Extract the residuals
    garch_residuals <- residuals(garch_fit)
    
    # Return the residuals of the GARCH(1,1) model
    return(as.numeric(garch_residuals))
  }
}


# List of data frames for each sector
sectors_log_returns <- list(
  communication = communication_period,
  discretionary = discretionary_period,
  staples = staples_period
)

# Initialize an empty data frame to store the GARCH results
garch_results <- data.frame(
  date = as.Date(character()),
  sector = character(),
  symbol = character(),
  residual = numeric(),
  passed_garch_test = logical()
)

# Loop through each sector
for (sector_name in names(sectors_log_returns)) {
  
  # Get the data for the current sector
  sector_data <- sectors_log_returns[[sector_name]]
  
  # Get the unique stock symbols in this sector
  stock_symbols <- unique(sector_data$symbol)
  
  # Loop through each stock
  for (stock in stock_symbols) {
    
    # Extract the log-returns and dates for the current stock
    stock_data <- sector_data[sector_data$symbol == stock, ]
    dates <- stock_data$date  # Assuming 'date' column exists
    log_returns <- stock_data$close_lgret
    
    # Run the GARCH test and fit function
    garch_result <- test_and_fit_garch(log_returns)
    
    # Determine if the GARCH model was fitted
    passed_garch_test <- !identical(log_returns, garch_result)
    
    # Add each entry as a separate row to the garch_results data frame
    for (i in seq_along(dates)) {
      garch_results <- rbind(
        garch_results,
        data.frame(
          date = dates[i],
          sector = sector_name,
          symbol = stock,
          residual = garch_result[i],
          passed_garch_test = passed_garch_test
        )
      )
    }
  }
}

# head(garch_results) 

# META, NFLX, TGT are the only symbols that did not pass the GARCH Test, because of this the data frame, 'garch_results' will keep the log returns for those symbols.


```


***Example of Using Garch Function Above***
```{r}
# Let's simulate some data from a GARCH(1, 1) and check our function

# Set the GARCH(1,1) model specifications with starting parameter values
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "norm", 
  fixed.pars = list(mu = 0, omega = 0.1, alpha1 = 0.1, beta1 = 0.8) # Specify parameter values
)


# Simulate n = 1000 data points from the GARCH(1,1) model
n <- 1000
x <- as.numeric(fitted(ugarchpath(spec, n.sim = n)))

# Plot the simulated GARCH data
plot(x, type = "l", col = "blue", main = "Simulated GARCH(1,1) Time Series", 
     xlab = "Time", ylab = "Value")

# Checking our function
fit_x <- test_and_fit_garch(x)

# Now, let's simulate white noise and re-check our function
y <- rnorm(n, mean = 0, sd = 1)

# Plot the simulated GARCH data
plot(y, type = "l", col = "blue", main = "Simulated Random Walk Time Series", 
     xlab = "Time", ylab = "Value")

# Re-checking our function
fit_y <- test_and_fit_garch(y)

```


***Now that we got 'garch_results', we are going to filtering everything out again into its sectors to we can then perform analysis for each stock sector***
```{r}

# Create separate data frames for each sector
garch_communication <- garch_results |> filter(sector == "communication")
garch_discretionary <- garch_results |> filter(sector == "discretionary")
garch_staples <- garch_results |> filter(sector == "staples")


# Now we are going to separate each dataframe further for stable, and pandemic periods

# Filtering for the communication services sector
garch_communication_stable <- garch_communication |>
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_communication_pandemic <- garch_communication |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")

# Filtering for the consumer discretionary sector
garch_discretionary_stable <- garch_discretionary |> 
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_discretionary_pandemic <- garch_discretionary |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")

# Filtering for the consumer staples sector
garch_staples_stable <- garch_staples |> 
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_staples_pandemic <- garch_staples |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")


```

***function to turn GARCH data frames to wide format***
```{r}
get_wide_residuals <- function(residuals) {
  wide_residuals <- residuals |> 
    select(date, symbol, residual) |> 
    pivot_wider(names_from = symbol, values_from = residual) |> 
    select(-date)
  
  return(wide_residuals)
}

```

***Wide format for GARCH dataframes***
```{r}

# Wide format for the communication services 
garch_communication_stable_wide <- get_wide_residuals(garch_communication_stable)

garch_communication_pandemic_wide <- get_wide_residuals(garch_communication_pandemic)

# Wide format for the consumer discretionary 
garch_discretionary_stable_wide <- get_wide_residuals(garch_discretionary_stable)

garch_discretionary_pandemic_wide <- get_wide_residuals(garch_discretionary_pandemic)

# Wide format for the consumer staples
garch_staples_stable_wide <- get_wide_residuals(garch_staples_stable)

garch_staples_pandemic_wide <- get_wide_residuals(garch_staples_pandemic)


```


***Creating Matrix Plots for Communication Services***
```{r}


# Helper function to reshape a matrix into a long format for ggplot, keeping only the lower triangle
reshape_for_ggplot <- function(matrix, names) {
  as.data.frame(matrix) %>%
    rownames_to_column(var = "Var1") %>%
    pivot_longer(-Var1, names_to = "Var2", values_to = "value") %>%
    mutate(
      Var1 = factor(Var1, levels = names, ordered = TRUE),
      Var2 = factor(Var2, levels = names, ordered = TRUE)
    ) %>%
    filter(as.numeric(Var1) > as.numeric(Var2))  # Keep only lower triangle
}

# Function to plot the Spearman Correlation
plot_spearman_correlation_ggplot <- function(wide_data, title, tickers) {
  # Calculate Spearman correlation matrix
  spearman_corr <- cor(wide_data, method = "spearman", use = "pairwise.complete.obs")
  
  # Reshape for ggplot
  corr_data <- reshape_for_ggplot(spearman_corr, tickers)
  
  # Plot with values inside the tiles
  ggplot(corr_data, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 3) +  # Display correlation values
    scale_fill_viridis_c(option = "plasma", limits = c(-1, 1), name = "Spearman\nCorrelation") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(size = 10)
    ) +
    labs(title = title, x = "", y = "") +
    coord_fixed()
}

# Function to compute Mutual Information matrix for a wide-format data frame
compute_mutual_information_matrix <- function(wide_data) {
  discretized_data <- discretize(wide_data)
  n <- ncol(discretized_data)
  mi_matrix <- matrix(0, n, n)
  
  for (i in 1:n) {
    for (j in 1:n) {
      mi_matrix[i, j] <- mutinformation(discretized_data[, i], discretized_data[, j])
    }
  }
  
  colnames(mi_matrix) <- colnames(wide_data)
  rownames(mi_matrix) <- colnames(wide_data)
  
  return(mi_matrix)
}

# Plotting function for Mutual Information matrix
plot_mutual_information_matrix <- function(wide_data, title, tickers) {
  mi_matrix <- compute_mutual_information_matrix(wide_data)
  mi_data <- reshape_for_ggplot(mi_matrix, tickers)
  
  ggplot(mi_data, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 3) +  # Display MI values
    scale_fill_viridis_c(option = "plasma", name = "Mutual\nInformation") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(size = 10)
    ) +
    labs(title = title, x = "", y = "") +
    coord_fixed()
}

# Function to compute Tail Dependence Coefficient (TDC) matrix
compute_tdc_matrix <- function(data, percentile = 0.05) {
  n <- nrow(data)
  percentile_value <- quantile(data[[1]], probs = percentile)
  k <- sum(data[[1]] <= percentile_value)
  tdc_mat <- FRAPO::tdc(data, method = "EmpTC", lower = TRUE, k = k)
  return(tdc_mat)
}

# Function to plot Tail Dependence Coefficient (TDC) matrix with only the lower triangle
plot_tdc_mat <- function(data, title, tickers, percentile = 0.05) {
  tdc_mat <- compute_tdc_matrix(data, percentile = percentile)
  tdc_mat[upper.tri(tdc_mat)] <- NA
  tdc_long <- reshape_for_ggplot(tdc_mat, tickers)
  
  ggplot(tdc_long, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 3, na.rm = TRUE) +
    scale_fill_viridis_c(option = "plasma", limits = c(0, 1), name = "TDC") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(size = 10),
      legend.position = "right"
    ) +
    labs(title = title, x = "", y = "") +
    coord_fixed()
}

# Outputting the all of the plots for the communication services sector
plot_spearman_correlation_ggplot(garch_communication_stable_wide, "Spearman Correlation for Communication Services (Stable Period)", communication_tickers)
plot_spearman_correlation_ggplot(garch_communication_pandemic_wide, "Spearman Correlation for Communication Services (Pandemic Period)", communication_tickers)

plot_mutual_information_matrix(garch_communication_stable_wide, "Mutual Information for Communication Services (Stable Period)", communication_tickers)
plot_mutual_information_matrix(garch_communication_pandemic_wide, "Mutual Information for Communication Services (Pandemic Period)", communication_tickers)

plot_tdc_mat(garch_communication_stable_wide, "TDC for Communication Services (Stable Period)", communication_tickers)
plot_tdc_mat(garch_communication_pandemic_wide, "TDC for Communication Services (Pandemic Period)", communication_tickers)



```


***Plots for the Consumer Discretionary Sector***
```{r}
# Spearman Correlation for the Consumer Discretionary Sector

plot_spearman_correlation_ggplot(garch_discretionary_stable_wide, "Spearman Correlation for Consumer Discretionary (Stable Period)", discretionary_tickers)
plot_spearman_correlation_ggplot(garch_discretionary_pandemic_wide, "Spearman Correlation for Consumer Discretionary (Pandemic Period)", discretionary_tickers)

plot_mutual_information_matrix(garch_discretionary_stable_wide, "Mutual Information for Consumer Discretionary (Stable Period)", discretionary_tickers)
plot_mutual_information_matrix(garch_discretionary_pandemic_wide, "Mutual Information for Consumer Discretionary (Pandemic Period)", discretionary_tickers)

plot_tdc_mat(garch_discretionary_stable_wide, "TDC for Consumer Discretionary (Stable Period)", discretionary_tickers)
plot_tdc_mat(garch_discretionary_pandemic_wide, "TDC for Consumer Discretionary (Pandemic Period)", discretionary_tickers)


```


***Plots for the Consumer Staples Sector***
```{r}

# Spearman Correlation for the Consumer Staples Sector

plot_spearman_correlation_ggplot(garch_staples_stable_wide, "Spearman Correlation for Consumer Staples (Stable Period)", staples_tickers)
plot_spearman_correlation_ggplot(garch_staples_pandemic_wide, "Spearman Correlation for Consumer Staples (Pandemic Period)", staples_tickers)

plot_mutual_information_matrix(garch_staples_stable_wide, "Mutual Information for Consumer Staples (Stable Period)", staples_tickers)
plot_mutual_information_matrix(garch_staples_pandemic_wide, "Mutual Information for Consumer Staples (Pandemic Period)", staples_tickers)

plot_tdc_mat(garch_staples_stable_wide, "TDC for Consumer Staples (Stable Period)", staples_tickers)
plot_tdc_mat(garch_staples_pandemic_wide, "TDC for Consumer Staples (Pandemic Period)", staples_tickers)

```


***Cross-Sector Analysis***
```{r}
# Below is a dataframe for the GARCH residuals for each sector, during the stable period
all_sector_garch_stable <- rbind(garch_communication_stable, garch_discretionary_stable, garch_staples_stable)

# Below is a dataframe for the GARCH residuals for each sector, during the pandemic period
all_sector_garch_pandemic <- rbind(garch_communication_pandemic, garch_discretionary_pandemic, garch_staples_pandemic)

all_sector_wide_stable <- get_wide_residuals(all_sector_garch_stable)
all_sector_wide_pandemic <- get_wide_residuals(all_sector_garch_pandemic)


all_tickers <- rbind(communication_tickers, discretionary_tickers, staples_tickers) # combining all of the Vectors that have the symbols for each stocl

plot_spearman_correlation_ggplot(all_sector_wide_stable, "Spearman Correlation Across All Sectors (Stable Period)", all_tickers)
plot_spearman_correlation_ggplot(all_sector_wide_pandemic, "Spearman Correlation for Across All Sectors (Pandemic Period)", all_tickers)

plot_mutual_information_matrix(all_sector_wide_stable, "Mutual Information Across All Sectors (Stable Period)", all_tickers)
plot_mutual_information_matrix(all_sector_wide_pandemic, "Mutual Information Across All Sectors (Pandemic Period)", all_tickers)

plot_tdc_mat(all_sector_wide_stable, "TDC Across All Sectors (Stable Period)", all_tickers)
plot_tdc_mat(all_sector_wide_pandemic, "TDC Across All Sectors (Pandemic Period)", all_tickers)


```