---
title: "Extreme Dependence Data Analysis"
author: "Alex Reyes"
date: "2024-10-08"
output: html_document
---

***Cleaning the Workspace/Loading Required libraries***
```{r}
# Clean the workspace
rm(list = ls())

getwd()

# Load required packages
library(tidyverse)
library(tidyselect)
library(tidyquant)
library(quantmod)
library(timetk)
library(copula)
library(VineCopula)
library(cowplot)
library(infotheo)
library(GGally)
library(nortest)
library(car)
library(knitr)
library(kableExtra)
library(pander)
library(patchwork)
library(rugarch)
library(FinTS) 
library(reshape2)
library(viridis)
library(tibble)
library(ggpubr)
```

***Functions for Stock Closing Prices/Log Returns/Wide-Format for Copula Analysis/plotting prices and log returns***
```{r}
# Create functions for tick marks
number_ticks <- function(n) { 
  function(limits) pretty(limits, n) 
}

date_ticks <- function(n) {
  function(limits) {
    pretty(seq(limits[1], limits[2], length.out = n), n)
  }
}

# Function to get and process stock prices
get_prices <- function(tickers, start_date, end_date, sector_name) {
  prices <- tq_get(tickers, from = start_date, to = end_date) |> 
    mutate(
      symbol = factor(symbol),
      sector = sector_name
    ) |> 
    group_by(symbol) |> 
    mutate(
      close_scaled = as.vector(scale(close)),  # Scale closing prices
      close_lgret = as.vector(log(close / lag(close)))  # Calculate log returns
    ) |> 
    ungroup() |> 
    drop_na() |> 
    select(date, sector, symbol, close, close_scaled, close_lgret)
  
  return(prices)
}

# Function to convert prices data to wide format for copula analysis
get_prices <- function(tickers, start_date, end_date, sector_name) {
  prices <- tq_get(tickers, from = start_date, to = end_date) %>%
    mutate(
      symbol = factor(symbol),
      sector = sector_name
    ) %>%
    group_by(symbol) %>%
    arrange(date, .by_group = TRUE) %>%  # Ensure data is sorted by date
    mutate(
      close_scaled = as.numeric(scale(close)),
      # Calculate log returns using the difference of logs; first value set to 0
      close_lgret = c(0, diff(log(close)))
    ) %>%
    ungroup() %>%
    select(date, sector, symbol, close, close_scaled, close_lgret)
  
  return(prices)
}


# Define start and end dates for the data
start_date <- "2006-12-31"
end_date <- "2024-01-01"

# Define tickers for the consumer discretionary sector
staples_tickers <- c(
  "COST", "WMT", "KO",
  "PEP", "TGT", "KR",
  "TSN", "CLX", "STZ",
  "MNST", "SYY", "PG")

# Define tickers for the communication services sector
communication_tickers <- c(
  "META", "GOOGL", "NFLX", 
  "DIS", "WBD", "PARA", 
  "CMCSA", "EA", "T", 
  "VZ", "TMUS", "OMC")

# Define tickers for the consumer staples sector
discretionary_tickers <- c(
  "AMZN", "TSLA", "HD",
  "MCD", "LOW", "NKE",
  "SBUX", "MAR", "EBAY",
  "BBY", "ETSY", "MGM")

technology_tickers <- c(
  "AAPL", "NVDA", "MSFT",
  "AVGO", "ORCL", "CRM", 
  "CSCO", "IBM", "ADBE", 
  "AMD", "HP", "CTSH")



## Fetch and process technology sector stock prices ##

tech_prices <- get_prices(
  tickers = technology_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Technology"
)

tech_global <- tech_prices |> 
  dplyr::filter(date >= "2007-01-01" & date <= "2009-12-31")
tech_global$period <- "Global Crisis"
tech_global <- tech_global |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

tech_stable <- tech_prices |> 
  dplyr::filter(date >= "2015-12-31" & date <= "2018-12-31")
tech_stable$period <- "Stable Market"
tech_stable <- tech_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

tech_pandemic <- tech_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
tech_pandemic$period <- "Pandemic and Downturn"
tech_pandemic <- tech_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods
tech_period <- rbind(
    tech_global, tech_stable, tech_pandemic
  ) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Global Crisis", "Stable Market", "Pandemic and Downturn")
    )
  )



## Fetch and process consumer staples sector stock prices ##
staples_prices <- get_prices(
  tickers = staples_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Consumer Staples"
)

staples_global <- staples_prices |> 
  dplyr::filter(date >= "2007-01-01" & date <= "2009-12-31")
staples_global$period <- "Global Crisis"
staples_global <- staples_global |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

staples_stable <- staples_prices |> 
  dplyr::filter(date >= "2015-12-31" & date <= "2018-12-31")
staples_stable$period <- "Stable Market"
staples_stable <- staples_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

staples_pandemic <- staples_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
staples_pandemic$period <- "Pandemic and Downturn"
staples_pandemic <- staples_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods
staples_period <- rbind(
    staples_global, staples_stable, staples_pandemic
  ) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Global Crisis", "Stable Market", "Pandemic and Downturn")
    )
  )


## Filtering and Processing stocks for Consumer Discretionary Sector ##

discretionary_prices <- get_prices(
  tickers = discretionary_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Consumer Discretionary"
)

discretionary_global <- discretionary_prices |> 
  dplyr::filter(date >= "2007-01-01" & date <= "2009-12-31")
discretionary_global$period <- "Global Crisis"
discretionary_global <- discretionary_global |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

discretionary_stable <- discretionary_prices |> 
  dplyr::filter(date >= "2015-12-31" & date <= "2018-12-31")
discretionary_stable$period <- "Stable Market"
discretionary_stable <- discretionary_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

discretionary_pandemic <- discretionary_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
discretionary_pandemic$period <- "Pandemic and Downturn"
discretionary_pandemic <- discretionary_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods
discretionary_period <- rbind(
    discretionary_global, discretionary_stable, discretionary_pandemic
  ) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Global Crisis", "Stable Market", "Pandemic and Downturn")
    )
  )


# Filtering and Processing stocks for Communication Services Sector

communication_prices <- get_prices(
  tickers = communication_tickers, 
  start_date = start_date, 
  end_date = end_date, 
  sector_name = "Communication Services"
)

communication_global <- communication_prices |> 
  filter(date >= "2007-01-01" & date <= "2009-12-31")
communication_global$period <- "Global Crisis"
communication_global <- communication_global |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

communication_stable <- communication_prices |> 
  filter(date >= "2015-12-31" & date <= "2018-12-31")
communication_stable$period <- "Stable Market"
communication_stable <- communication_stable |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

communication_pandemic <- communication_prices |> 
  dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")
communication_pandemic$period <- "Pandemic and Downturn"
communication_pandemic <- communication_pandemic |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  )

# Combine the periods
communication_period <- rbind(
    communication_global, communication_stable, communication_pandemic
  ) |> 
  select(
    date, period, sector, symbol, close, close_scaled, close_lgret
  ) |> 
  mutate(
    period = factor(
      period, 
      ordered = TRUE,
      levels = c("Global Crisis", "Stable Market", "Pandemic and Downturn")
    )
  )


```


***Plotting Scaled Closing Prices/Log Returns***
```{r}

# Function to plot scaled closing prices of stocks
plot_stocks <- function(prices, sector_name) {
  
  # Define the crisis periods
  crisis_periods <- data.frame(
    start = as.Date(
      c("2007-01-01", "2015-12-31", "2019-12-31")
    ),
    end = as.Date(
      c("2009-12-31", "2018-12-31", "2023-12-31")
    ),
    crisis = c(
      "Global Crisis", "Stable Market", "Pandemic and Downturn"
    )
  ) |> 
    mutate(crisis = factor(
      crisis, ordered = TRUE, 
        levels = c(
          "Global Crisis", "Stable Market", "Pandemic and Downturn"
        )
      )
    )
  
  
  stock_plot <- prices |> 
    ggplot(aes(x = date, y = close_scaled, color = symbol, group = symbol)) +
    geom_rect(
      data = crisis_periods, 
      aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf, fill = crisis),
      alpha = 0.45, inherit.aes = FALSE
    ) +
    geom_line(linewidth = 0.75) +
    scale_color_brewer(palette = "Set3") +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_fill_brewer(palette = "Pastel1") +
    labs(
      title = paste0("Scaled Closing Prices of ", sector_name, " Sector Stocks"),
      x = NULL,
      y = "Scaled Closing Prices",
      color = "Stock Symbol",
      fill = "Crisis Period"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(vjust = 1, size = 14, face = "bold"),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      legend.box = "vertical",
      legend.box.just = "center"
    ) + 
    guides(
      color = guide_legend(order = 1, nrow = 2, byrow = TRUE),
      fill = guide_legend(order = 2)
    )

  return(stock_plot)
}

# Function to plot log returns of the scaled closing prices of stocks for the study periods
plot_periods <- function(periods, sector_name) {

  periods |> 
    ggplot(
      aes(x = date, y = close_lgret, color = symbol, group = symbol)
    ) + 
    geom_line(linewidth = 0.75) +
    scale_color_brewer(palette = "Set3") +
    scale_x_date(breaks = date_ticks(5), date_labels = "%Y") +  
    scale_y_continuous(breaks = number_ticks(5)) +
    scale_fill_brewer(palette = "Pastel1") +
    facet_wrap(~period, scales = "free_x") +
    labs(
      title = paste0("Log Returns for the ", sector_name, " Sector Stocks"),
      x = NULL,
      y = "Log Returns of Closing Prices",
      color = "Stock Symbol"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(vjust = 1, size = 14, face = "bold"),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      legend.box = "vertical",
      legend.box.just = "center",
      plot.margin = margin(5.5, 30, 5.5, 5.5), 
      panel.spacing = unit(2, "lines") 
    ) + 
    guides(
      color = guide_legend(order = 1, nrow = 2, byrow = TRUE)
    ) -> plot_periods
  
  return(plot_periods)
  
}


plot_stocks(tech_prices, "Technology")


```


***Testing ARCH Effects, and Fitting one when necessary (significant P-value)***
```{r}


#' Test for ARCH effects and fit a GARCH(1,1) model if necessary
#'
#' This function takes a time series of log-returns (numerical vector) as input, 
#' performs an ARCH test using the FinTS::ArchTest function, and checks whether 
#' a GARCH structure is necessary. If the ARCH test indicates the need for a GARCH model, 
#' the function fits a GARCH(1,1) model using the rugarch package and returns the residuals.
#' If the ARCH test fails (i.e., no GARCH structure is necessary), it simply returns the input time series.
#'
#' @param x A numerical vector containing the time series of log-returns.
#' @param lags The number of lags for the ARCH test (default is 5).
#' @param alpha The significance level for the ARCH test (default is 0.05).
#' 
#' @return If no ARCH effects are found, the function returns the original time series. 
#' If ARCH effects are detected, it returns the residuals of the fitted GARCH(1,1) model.
#'
#' @examples
#' # Example with simulated GARCH(1,1) data:
#' n <- 1000
#' x <- as.numeric(fitted(ugarchpath(ugarchspec(), n.sim = n)))
#' result <- test_and_fit_garch(x)
#' print(result)
#'
#' @import rugarch
#' @import FinTS
#' @export
test_and_fit_garch <- function(x, lags = 5, alpha = 0.05) {
  
  # Step 1: Perform the ARCH test using FinTS::ArchTest
  arch_test <- FinTS::ArchTest(x, lags = lags)
  
  # Extract the p-value from the test
  p_value <- arch_test$p.value
  
  # Step 2: Decision rule based on p-value
  if (p_value >= alpha) {
    # If the p-value is greater than alpha, no ARCH effects are detected
    print("No significant ARCH effects. A GARCH model might not be necessary.")
    return(x)  # Return the original time series
  } else {
    # If the p-value is less than alpha, ARCH effects are detected
    print("There is evidence of ARCH effects. A GARCH model may be necessary.")
    
    # Step 3: Fit the GARCH(1,1) model to the data 'x'
    garch_spec <- ugarchspec(
      variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
      mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
      distribution.model = "norm"
    )
    
    # Fit the model
    garch_fit <- ugarchfit(spec = garch_spec, data = x)
    
    # Step 4: Extract the residuals
    garch_residuals <- residuals(garch_fit)
    
    # Return the residuals of the GARCH(1,1) model
    return(as.numeric(garch_residuals))
  }
}


# List of data frames for each sector
sectors_log_returns <- list(
  communication = communication_period,
  discretionary = discretionary_period,
  staples = staples_period,
  tech = tech_period
)

# Initialize an empty data frame to store the GARCH results
garch_results <- data.frame(
  date = as.Date(character()),
  sector = character(),
  symbol = character(),
  residual = numeric(),
  passed_garch_test = logical()
)

# Loop through each sector
for (sector_name in names(sectors_log_returns)) {
  
  # Get the data for the current sector
  sector_data <- sectors_log_returns[[sector_name]]
  
  # Get the unique stock symbols in this sector
  stock_symbols <- unique(sector_data$symbol)
  
  # Loop through each stock
  for (stock in stock_symbols) {
    
    # Extract the log-returns and dates for the current stock
    stock_data <- sector_data[sector_data$symbol == stock, ]
    dates <- stock_data$date  # Assuming 'date' column exists
    log_returns <- stock_data$close_lgret
    
    # Run the GARCH test and fit function
    garch_result <- test_and_fit_garch(log_returns)
    
    # Determine if the GARCH model was fitted
    passed_garch_test <- !identical(log_returns, garch_result)
    
    # Add each entry as a separate row to the garch_results data frame
    for (i in seq_along(dates)) {
      garch_results <- rbind(
        garch_results,
        data.frame(
          date = dates[i],
          sector = sector_name,
          symbol = stock,
          residual = garch_result[i],
          passed_garch_test = passed_garch_test
        )
      )
    }
  }
}

# head(garch_results) 

# META, NFLX, TGT are the only symbols that did not pass the GARCH Test, because of this the data frame, 'garch_results' will keep the log returns for those symbols.


```


***Example of Using Garch Function Above***
```{r}
# Let's simulate some data from a GARCH(1, 1) and check our function

# Set the GARCH(1,1) model specifications with starting parameter values
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "norm", 
  fixed.pars = list(mu = 0, omega = 0.1, alpha1 = 0.1, beta1 = 0.8) # Specify parameter values
)


# Simulate n = 1000 data points from the GARCH(1,1) model
n <- 1000
x <- as.numeric(fitted(ugarchpath(spec, n.sim = n)))

# Plot the simulated GARCH data
plot(x, type = "l", col = "blue", main = "Simulated GARCH(1,1) Time Series", 
     xlab = "Time", ylab = "Value")

# Checking our function
fit_x <- test_and_fit_garch(x)

# Now, let's simulate white noise and re-check our function
y <- rnorm(n, mean = 0, sd = 1)

# Plot the simulated GARCH data
plot(y, type = "l", col = "blue", main = "Simulated Random Walk Time Series", 
     xlab = "Time", ylab = "Value")

# Re-checking our function
fit_y <- test_and_fit_garch(y)

```


***Now that we got 'garch_results', we are going to filtering everything out again into its sectors to we can then perform analysis for each stock sector***
```{r}

# Create separate data frames for each sector
garch_communication <- garch_results |> dplyr::filter(sector == "communication")
garch_discretionary <- garch_results |> dplyr::filter(sector == "discretionary")
garch_staples <- garch_results |> dplyr::filter(sector == "staples")
garch_tech <- garch_results |> dplyr::filter(sector == "technology")

# Now we are going to separate each dataframe further for stable, and pandemic periods

# Filtering for the communication services sector
garch_communication_stable <- garch_communication |>
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_communication_pandemic <- garch_communication |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")

# Filtering for the consumer discretionary sector
garch_discretionary_stable <- garch_discretionary |> 
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_discretionary_pandemic <- garch_discretionary |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")

# Filtering for the consumer staples sector
garch_staples_stable <- garch_staples |> 
    dplyr::filter(date >= "2016-01-01" & date <= "2018-12-31")
  
garch_staples_pandemic <- garch_staples |> 
    dplyr::filter(date >= "2019-12-31" & date <= "2023-12-31")


```

***function to turn GARCH data frames to wide format***
```{r}
get_wide_residuals <- function(residuals) {
  wide_residuals <- residuals |> 
    select(date, symbol, residual) |> 
    pivot_wider(names_from = symbol, values_from = residual) |> 
    select(-date)
  
  return(wide_residuals)
}

```

***Wide format for GARCH dataframes***
```{r}

# Wide format for the communication services 
garch_communication_stable_wide <- get_wide_residuals(garch_communication_stable)

garch_communication_pandemic_wide <- get_wide_residuals(garch_communication_pandemic)

# Wide format for the consumer discretionary 
garch_discretionary_stable_wide <- get_wide_residuals(garch_discretionary_stable)

garch_discretionary_pandemic_wide <- get_wide_residuals(garch_discretionary_pandemic)

# Wide format for the consumer staples
garch_staples_stable_wide <- get_wide_residuals(garch_staples_stable)

garch_staples_pandemic_wide <- get_wide_residuals(garch_staples_pandemic)

r_data <- list(garch_communication_stable_wide, garch_communication_pandemic_wide, garch_discretionary_stable_wide, garch_discretionary_pandemic_wide, garch_staples_stable_wide, garch_staples_pandemic_wide)



```


***Plots for Spearman Correlation and Mutual Information***
```{r}

# Helper function to reshape a matrix into a long format for ggplot, keeping only the lower triangle
reshape_for_ggplot <- function(matrix, name_levels) {
  as.data.frame(matrix) %>%
    tibble::rownames_to_column(var = "Var1") %>%
    tidyr::pivot_longer(cols = -Var1, 
                        names_to = "Var2", 
                        values_to = "value") %>%
    dplyr::mutate(
      Var1 = factor(Var1, levels = name_levels, ordered = TRUE),
      Var2 = factor(Var2, levels = name_levels, ordered = TRUE)
    ) %>%
    dplyr::filter(as.numeric(Var1) > as.numeric(Var2))
}


# Function to plot the Spearman Correlation
plot_spearman_correlation_ggplot <-function(wide_data, title, subtitle, tickers) {
  # Calculate Spearman correlation matrix
  spearman_corr <- cor(wide_data, method = "spearman", use = "pairwise.complete.obs")
  
  # Reshape for ggplot
  corr_data <- reshape_for_ggplot(spearman_corr, tickers)
  
  # Plot with values inside the tiles
  ggplot(corr_data, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") + # Edit the grid lines
    geom_text(aes(label = round(value, 2)), color = "black", size = 8) +  # Display correlation values
    scale_fill_viridis_c(option = "plasma", limits = c(-1, 1)) +
    ggtitle(title, subtitle) +
    labs(x = "", y = "", fill = NULL) +
    coord_fixed() +
    theme_minimal() +
    guides(fill = guide_colorbar(barwidth = 30, barheight = 1),
           label.theme = element_text(size = 6))+
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      plot.title.position = "plot",
      plot.title = element_text(size = 30, face = "bold"),
      plot.subtitle = element_text(size = 24),
      plot.margin = grid::unit(c(0, 0, 0, 0), "lines"), 
      legend.text = element_text(size = 10)
    )
}


# Function to compute Mutual Information matrix for a wide-format data frame
compute_mutual_information_matrix <- function(wide_data) {
  discretized_data <- discretize(wide_data)
  n <- ncol(discretized_data)
  mi_matrix <- matrix(0, n, n)
  
  for (i in 1:n) {
    for (j in 1:n) {
      mi_matrix[i, j] <- mutinformation(discretized_data[, i], discretized_data[, j])
    }
  }
  

  colnames(mi_matrix) <- colnames(wide_data)
  rownames(mi_matrix) <- colnames(wide_data)
  
  return(mi_matrix)
}


# Getting the max MI of three sectors

# 1. Compute MI matrices for each sector
mi_comm_stable_score<- compute_mutual_information_matrix(garch_communication_stable_wide)
mi_comm_pan_score<-compute_mutual_information_matrix(garch_communication_pandemic_wide)

mi_dis_stable_score<- compute_mutual_information_matrix(garch_discretionary_stable_wide)
mi_dis_pan_score<- compute_mutual_information_matrix(garch_discretionary_pandemic_wide)

mi_staples_stable_score<- compute_mutual_information_matrix(garch_staples_stable_wide)
mi_staples_pan_score<- compute_mutual_information_matrix(garch_staples_pandemic_wide)
# 2. Get the overall maximum MI across all sectors
max_mi <- max(mi_comm_stable_score,mi_comm_pan_score, mi_dis_stable_score, mi_dis_pan_score, mi_staples_stable_score, mi_staples_pan_score,na.rm = TRUE)



# Plotting function for Mutual Information matrix
plot_mutual_information_matrix <- function(wide_data, title, subtitle, tickers) {
  mi_matrix <- compute_mutual_information_matrix(wide_data)
  mi_data <- reshape_for_ggplot(mi_matrix, tickers)
  
    # add a normalized column
  mi_data$norm_value <- mi_data$value / max_mi
  
  ggplot(mi_data, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 8) +  # Display MI values
    scale_fill_viridis_c(option = "plasma", limits= c(0, 1),name = "") +
    ggtitle(title, subtitle) +
    theme_minimal() +
    guides(fill = guide_colorbar(barwidth = 30, barheight = 1), 
           label.theme = element_text(size = 6))+
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      plot.subtitle = element_text(size = 24), 
      plot.title = element_text(size = 30, face = "bold"),
      legend.text = element_text(size = 10),
    ) +
    
    labs(title = title, subtitle = subtitle, x = "", y = "") +
    coord_fixed()
}










mi_stable_comm <-plot_mutual_information_matrix(garch_communication_stable_wide, "","", communication_tickers)
mi_pandemic_comm <-plot_mutual_information_matrix(garch_communication_pandemic_wide, "","", communication_tickers)





mi_stable_dis <-plot_mutual_information_matrix(garch_discretionary_stable_wide, "","", discretionary_tickers)
mi_pan_dis <-mutualinfo_pan_dis <-plot_mutual_information_matrix(garch_discretionary_pandemic_wide, "", "",discretionary_tickers)




mi_stable_con <-plot_mutual_information_matrix(garch_staples_stable_wide, "", "",staples_tickers)
mi_pan_con <-plot_mutual_information_matrix(garch_staples_pandemic_wide, "", "",staples_tickers)



```


***TDC Computations and Plotting***
```{r}
## Function to compute pairwise TDC using vineCopula for each pair of stocks.
compute_tdc_matrix <- function(data, percentile = 0.05, families = NA, rotations = TRUE) {
  # data: a data frame or matrix with each column representing a stock's time series.
  # percentile: the q value at which to approximate the lower TDC.
  # families & rotations: passed on to BiCopSelect for copula selection.
  
  # Get ticker names from column names (if available) and number of series.
  tickers <- colnames(data)
  n_stocks <- ncol(data)
  
  # Initialize an empty matrix to store the approximated TDC values.
  tdc_mat <- matrix(NA, nrow = n_stocks, ncol = n_stocks,
                    dimnames = list(tickers, tickers))
  
  # Loop over all unique pairs of stocks.
  for (i in 1:(n_stocks - 1)) {
    for (j in (i + 1):n_stocks) {
      # Extract the two series.
      u <- data[[i]]
      v <- data[[j]]
      
      # If data are not already pseudo-observations on [0,1], convert them.
      # (Simple conversion: rank / (n+1))
      if (min(u) < 0 || max(u) > 1) {
        u <- rank(u) / (length(u) + 1)
      }
      if (min(v) < 0 || max(v) > 1) {
        v <- rank(v) / (length(v) + 1)
      }
      
      # 1. Fit the best copula for (u,v) using vineCopula's BiCopSelect.
      best_cop <- BiCopSelect(u, v, familyset = families, rotations = rotations)
      
      # 2. Compute the copula CDF at (q,q) for the selected model.
      C_val <- BiCopCDF(percentile, percentile,
                        family = best_cop$family,
                        par    = best_cop$par,
                        par2   = best_cop$par2)
      
      # 3. Approximate the lower tail dependence coefficient.
      approx_lower_tdc <- C_val / percentile
      
      # Store the result in the (i,j) and (j,i) positions (since TDC is symmetric).
      tdc_mat[i, j] <- approx_lower_tdc
      tdc_mat[j, i] <- approx_lower_tdc
    }
  }
  
  # Optionally, set the diagonal to 1 (a variable is perfectly tail–dependent with itself).
  diag(tdc_mat) <- 1
  
  return(tdc_mat)
}


#  mymat <- compute_tdc_matrix(garch_communication_stable_wide)
#  tdc_long <- reshape_for_tdc(mymat, tickers = communication_tickers)
## Helper function to reshape a square matrix for ggplot

# Reshape helper for TDC matrix
reshape_for_tdc <- function(mat, tickers) {
  # Uses melt from reshape2 to convert the matrix into long format.
  # If tickers are provided, they will be used as the row/column names.
 # Reshape the matrix into long format

  # Convert the matrix to long format
  mat_long <- as.data.frame(mat) %>%
    tibble::rownames_to_column(var = "Var1") %>%
    tidyr::pivot_longer(
      cols = -Var1,
      names_to = "Var2",
      values_to = "value"
    ) %>%
    dplyr::mutate(
      Var1 = factor(Var1, levels = tickers, ordered = TRUE),
      Var2 = factor(Var2, levels = tickers, ordered = TRUE)
    ) %>%
    dplyr::filter(as.numeric(Var1) > as.numeric(Var2))  # strictly lower triangle

  return(mat_long)
}


## Function to plot the TDC matrix (only the lower triangle)
plot_tdc_mat <- function(data, title, subtitle ,tickers = NULL, percentile = 0.05,
                         families = NA, rotations = TRUE) {
  # Compute the TDC matrix using the vineCopula approach.
  tdc_mat <- compute_tdc_matrix(data, percentile = percentile,
                                families = families, rotations = rotations)
  
  
  # Reshape the matrix into a long-format data frame.
  tdc_long <- reshape_for_tdc(tdc_mat, tickers)
  
  # Create the heatmap with ggplot2.
  p <- ggplot(tdc_long, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = ifelse(is.na(value), "", round(value, 2))),
              color = "black", size = 8, na.rm = TRUE) +
    scale_fill_viridis_c(option = "plasma", limits = c(0, 1),
                         name = "") +
      guides(fill = guide_colorbar(barwidth = 30, barheight = 1), 
           label.theme = element_text(size = 8))+
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      plot.subtitle = element_text(size = 24),
      plot.title = element_text(size = 30, face = "bold"),
      legend.text = element_text(size = 10),
      legend.position = "bottom"
    ) +
    labs(title = title,subtitle = subtitle ,x = "", y = "") +
    coord_fixed()
  
  return(p)
}



tdc_stable_comm <-plot_tdc_mat(garch_communication_stable_wide, "TDC for the Communication Services Sector (Stable Period)", " ",tickers =  communication_tickers, percentile = 0.05)
tdc_pan_comm <-plot_tdc_mat(garch_communication_pandemic_wide, "TDC for the Communication Services Sector (Pandemic Period)", " ",tickers =  communication_tickers, percentile = 0.05)



## Consumer Staples TDC ##
tdc_stable_con <-plot_tdc_mat(garch_staples_stable_wide, "TDC for the Consumer Staples Sector (Stable Period)", " ", tickers =  staples_tickers, percentile = 0.05)
tdc_pan_con <-plot_tdc_mat(garch_staples_pandemic_wide, "TDC for the Consumer Staples Sector (Pandemic Period)", " ", tickers = staples_tickers, percentile = 0.05)



## Consumer Discretionary TDC ##
tdc_stable_dis <-plot_tdc_mat(garch_discretionary_stable_wide, "TDC for the Consumer Discretionary Sector (Stable Period)", " ",tickers =  discretionary_tickers, percentile = 0.05)
tdc_pan_dis <-plot_tdc_mat(garch_discretionary_pandemic_wide, "TDC for the Consumer Discretionary Sector (Pandemic Period)", " ", tickers =  discretionary_tickers, percentile = 0.05)



```

***Cowplot Mutual Information***
```{r}


# Outputting the all of the plots for the communication services sector
mi_stable_comm <- plot_mutual_information_matrix(
  garch_communication_stable_wide,
  "                   Stable Period",
  "Communication Services", 
  communication_tickers
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(mi_stable_comm, position = NULL)
print(legend)

mi_stable_comm <- mi_stable_comm + theme(legend.position = "none")

mi_pan_comm <- plot_mutual_information_matrix(
  garch_communication_pandemic_wide,
  "               Pandemic Period",  
  "", 
  communication_tickers
) + theme(legend.position = "none")


comm_row <- plot_grid(
  mi_stable_comm, mi_pan_comm, nrow = 1
)


mi_stable_dis <- plot_mutual_information_matrix(
  garch_discretionary_stable_wide, 
  "",
  "Consumer Discretionary", 
  discretionary_tickers
) + theme(legend.position = "none")

mi_pan_dis <- plot_mutual_information_matrix(
  garch_discretionary_pandemic_wide, 
  "", 
  "",
  discretionary_tickers
) + theme(legend.position = "none")

dis_row <- plot_grid(
  mi_stable_dis, mi_pan_dis, nrow = 1
)

mi_stable_con <- plot_mutual_information_matrix(
  garch_staples_stable_wide, 
  "",
  "Consumer Staples", 
  staples_tickers
) + theme(legend.position = "none")

mi_pan_con <- plot_mutual_information_matrix(
  garch_staples_pandemic_wide, 
  "",
  "",
  staples_tickers
) + theme(legend.position = "none")

con_row <- plot_grid(
  mi_stable_con, mi_pan_con, nrow = 1
)


main_title <- ggdraw() 


combined_plot <-plot_grid(
  main_title,
  comm_row, dis_row, con_row, my_legend,
  nrow = 5,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)


save_plot("combined_plot_mi.png", 
          combined_plot, 
          base_width = 20,   # in inches
          base_height = 30,  # in inches
          dpi = 400)

```


***Cowplot Spearman Correlation***
```{r}


# Outputting the all of the plots for the communication services sector
Spearman_stable_comm <- plot_spearman_correlation_ggplot(
  garch_communication_stable_wide,
  "                     Stable Period",
  "Communication Services", 
  communication_tickers
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(Spearman_stable_comm, position = NULL)
print(legend)

Spearman_stable_comm <- Spearman_stable_comm + theme(legend.position = "none")

Spearman_pan_comm <- plot_spearman_correlation_ggplot(
  garch_communication_pandemic_wide,
  "                        Pandemic Period",  
  "", 
  communication_tickers
) + theme(legend.position = "none")


comm_row <- plot_grid(
  Spearman_stable_comm, Spearman_pan_comm, nrow = 1
)


Spearman_stable_dis <- plot_spearman_correlation_ggplot(
  garch_discretionary_stable_wide, 
  "",
  "Consumer Discretionary", 
  discretionary_tickers
) + theme(legend.position = "none")

Spearman_pan_dis <- plot_spearman_correlation_ggplot(
  garch_discretionary_pandemic_wide, 
  "", 
  "",
  discretionary_tickers
) + theme(legend.position = "none")

dis_row <- plot_grid(
  Spearman_stable_dis, Spearman_pan_dis, nrow = 1
)

Spearman_stable_con <- plot_spearman_correlation_ggplot(
  garch_staples_stable_wide, 
  "",
  "Consumer Staples", 
  staples_tickers
) + theme(legend.position = "none")

Spearman_pan_con <- plot_spearman_correlation_ggplot(
  garch_staples_pandemic_wide, 
  "",
  "",
  staples_tickers
) + theme(legend.position = "none")

con_row <- plot_grid(
  Spearman_stable_con, Spearman_pan_con, nrow = 1
)


main_title <- ggdraw() 


combined_plot <-plot_grid(
  main_title,
  comm_row, dis_row, con_row, my_legend,
  nrow = 5,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)


save_plot("combined_plot.png", 
          combined_plot, 
          base_width = 20,   # in inches
          base_height = 30,  # in inches
          dpi = 400)


```




***Cowplot TDC***
```{r}
# Outputting the all of the plots for the communication services sector
tdc_stable_comm <- plot_tdc_mat(
  garch_communication_stable_wide,
  "                   Stable Period",
  "Communication Services", 
  communication_tickers, 
  percentile = 0.05
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(tdc_stable_comm, position = NULL)
print(legend)

tdc_stable_comm <- tdc_stable_comm + theme(legend.position = "none")

tdc_pan_comm <- plot_tdc_mat(
  garch_communication_pandemic_wide,
  "               Pandemic Period",  
  "", 
  tickers = communication_tickers, 
  percentile = 0.05
) + theme(legend.position = "none")


comm_row <- plot_grid(
  tdc_stable_comm, tdc_pan_comm, nrow = 1
)


tdc_stable_dis <- plot_tdc_mat(
  garch_discretionary_stable_wide, 
  "",
  "Consumer Discretionary", 
 tickers = discretionary_tickers, 
  percentile = 0.05
) + theme(legend.position = "none")

tdc_pan_dis <- plot_tdc_mat(
  garch_discretionary_pandemic_wide, 
  "", 
  "",
  tickers = discretionary_tickers,
  percentile = 0.05
) + theme(legend.position = "none")

dis_row <- plot_grid(
  tdc_stable_dis, tdc_pan_dis, nrow = 1
)

tdc_stable_con <- plot_tdc_mat(
  garch_staples_stable_wide, 
  "",
  "Consumer Staples", 
 tickers = staples_tickers, 
  percentile = 0.05
) + theme(legend.position = "none")

tdc_pan_con <- plot_tdc_mat(
  garch_staples_pandemic_wide, 
  "",
  "",
 tickers = staples_tickers, 
  percentile = 0.05
) + theme(legend.position = "none")

con_row <- plot_grid(
  tdc_stable_con, tdc_pan_con, nrow = 1
)


main_title <- ggdraw()

combined_plot <-plot_grid(
  main_title,
  comm_row, dis_row, con_row, my_legend,
  nrow = 5,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)


save_plot("combined_plot_tdc.png", 
          combined_plot, 
          base_width = 20,   # in inches
          base_height = 30,  # in inches
          dpi = 400)

```



***These functions are to plot all sectors***
```{r}


# Load necessary libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(viridis)

# Revised helper: Reshape a square matrix into a long data frame for ggplot2,
# retaining only the strictly lower triangle (optionally excluding the diagonal).
reshape_for_ggplot_lower_triangle <- function(mat, ticker_names, remove_diagonal = TRUE) {
  # Ensure the matrix has the ticker names both as row and column names.
  rownames(mat) <- ticker_names
  colnames(mat) <- ticker_names
  
  # Convert the matrix to a long data frame.
  df <- as.data.frame(mat) %>%
    tibble::rownames_to_column("row_name") %>%
    pivot_longer(
      cols = -row_name,
      names_to = "col_name",
      values_to = "value"
    ) %>%
    mutate(
      # Save numeric indices so that we know where each ticker is on the axes.
      row_idx = match(row_name, ticker_names),
      col_idx = match(col_name, ticker_names)
    )
  
  # For the lower triangle, keep rows where row index > column index.
  if (remove_diagonal) {
    df <- df %>% filter(row_idx > col_idx)
  } else {
    df <- df %>% filter(row_idx >= col_idx)
  }
  
  return(df)
}

# Function to plot Spearman Correlation (lower triangle with vertical separators)
plot_spearman_correlation_ggplot_all_sector <- function(wide_data, title, tickers) {
  # Calculate the Spearman correlation matrix.
  spearman_corr <- cor(wide_data, method = "spearman", use = "pairwise.complete.obs")
  
  # Reshape the correlation matrix.
  corr_data <- reshape_for_ggplot_lower_triangle(spearman_corr, tickers, remove_diagonal = TRUE)
  
  n <- length(tickers)
  
  # Define vertical separator positions.
  vertical_lines <- tibble(
    x = seq(12.5, n - 0.5, by = 12)
  )
  
  ggplot(corr_data, aes(x = col_idx, y = row_idx, fill = value)) +
    geom_tile(color = "white", size = 0.1) +
    # Draw vertical separators with swapped endpoints.
    geom_segment(
      data = vertical_lines,
      aes(x = x, xend = x, y = 0, yend = n + 0.5),
      inherit.aes = FALSE,
      color = "black", size = 0.8
    ) +
    scale_fill_viridis_c(option = "plasma", limits = c(-1, 1), name = "") +
    scale_x_continuous(breaks = 1:n, labels = tickers) +
    scale_y_reverse(breaks = 1:n, labels = tickers) +  # Reverse y-axis so row 1 is at the top.
    coord_fixed() +
    theme_classic() +
    theme(
      axis.text.x  = element_text(angle = 45, vjust = 1, hjust = 1, size = 8, face = "bold"),
      axis.text.y  = element_text(size = 8, face = "bold"),
      axis.title   = element_blank(),
      panel.grid   = element_blank(),
      plot.title   = element_text(size = 25, hjust = 0.5),
      legend.title = element_text(size = 10),
      legend.text  = element_text(size = 6),
      legend.key.height = unit(0.3, "cm"),
      legend.key.width  = unit(0.15, "cm"),
      legend.position   = "bottom",
      plot.margin       = margin(0, 0, 0, 0)
    ) +
    guides(fill = guide_colorbar(barwidth = 25, barheight = 1)) +
    labs(title = title)
}

# Function to plot Tail Dependence Coefficient (TDC) matrix (lower triangle with vertical separators)
plot_tdc_mat_all_sector <- function(data, title, tickers, percentile = 0.05,
                                    families = NA, rotations = TRUE) {
  # Compute the TDC matrix.
  tdc_mat <- compute_tdc_matrix(data, percentile = percentile,
                                families = families, rotations = rotations)
  
  # Reshape the TDC matrix.
  tdc_data <- reshape_for_ggplot_lower_triangle(tdc_mat, tickers, remove_diagonal = TRUE)
  
  n <- length(tickers)
  
  vertical_lines <- tibble(
    x = seq(12.5, n - 0.5, by = 12)
  )
  
  ggplot(tdc_data, aes(x = col_idx, y = row_idx, fill = value)) +
    geom_tile(color = "white", size = 0.1) +
    geom_segment(
      data = vertical_lines,
      aes(x = x, xend = x, y = 0, yend = n + 0.5),
      inherit.aes = FALSE,
      color = "black", size = 0.8
    ) +
    scale_fill_viridis_c(option = "plasma", limits = c(0, 1), name = "") +
    scale_x_continuous(breaks = 1:n, labels = tickers) +
    scale_y_reverse(breaks = 1:n, labels = tickers) +
    coord_fixed() +
    theme_classic() +
    theme(
      axis.text.x  = element_text(angle = 45, vjust = 1, hjust = 1, size = 8, face = "bold"),
      axis.text.y  = element_text(size = 8, face = "bold"),
      axis.title   = element_blank(),
      panel.grid   = element_blank(),
      plot.title   = element_text(size = 25, hjust = 0.5),
      legend.title = element_text(size = 10),
      legend.text  = element_text(size = 6),
      legend.key.height = unit(0.3, "cm"),
      legend.key.width  = unit(0.15, "cm"),
      legend.position   = "bottom",
      plot.margin       = margin(0, 0, 0, 0)
    ) +
    guides(fill = guide_colorbar(barwidth = 25, barheight = 1)) +
    labs(title = title)
}

# returns a list with the normalized matrix and the max MI
compute_normalized_mi <- function(data) {
  # raw MI matrix
  mi_mat <- compute_mutual_information_matrix(data)
  # highest *pairwise* MI (exclude diagonal)
  max_mi <- max(mi_mat[lower.tri(mi_mat)], na.rm = TRUE)
  # normalize everything to [0,1]
  mi_mat_norm <- mi_mat / max_mi
  list(
    mi_mat = mi_mat_norm,
    max_mi = max_mi
  )
}


# Function to plot Mutual Information (MI) matrix (lower triangle with vertical separators)
plot_mi_mat_all_sector <- function(data, title, tickers, ...) {
    # 1. raw MI
# 1. get normalized MI + record raw max
  res      <- compute_normalized_mi(data)
  mi_mat   <- res$mi_mat       # already scaled to [0,1]
  max_mi   <- res$max_mi       # raw maximum
  
  # 4. reshape only lower triangle
  mi_data <- reshape_for_ggplot_lower_triangle(mi_mat, tickers, remove_diagonal = TRUE)
  
  n <- length(tickers)
  
  vertical_lines <- tibble(
    x = seq(12.5, n - 0.5, by = 12)
  )
  
  # Set the maximum MI value for the fill scale.
  max_mi <- max(mi_data$value, na.rm = TRUE)
  
  ggplot(mi_data, aes(x = col_idx, y = row_idx, fill = value)) +
    geom_tile(color = "white", size = 0.1) +
    geom_segment(
      data = vertical_lines,
      aes(x = x, xend = x, y = 0, yend = n + 0.5),
      inherit.aes = FALSE,
      color = "black", size = 0.8
    ) +
    scale_fill_viridis_c(option = "plasma", limits = c(0, 1), name = "") +
    scale_x_continuous(breaks = 1:n, labels = tickers) +
    scale_y_reverse(breaks = 1:n, labels = tickers) +
    coord_fixed() +
    theme_classic() +
    theme(
      axis.text.x  = element_text(angle = 45, vjust = 1, hjust = 1, size = 8, face = "bold"),
      axis.text.y  = element_text(size = 8, face = "bold"),
      axis.title   = element_blank(),
      panel.grid   = element_blank(),
      plot.title   = element_text(size = 25, hjust = 0.5),
      legend.title = element_text(size = 10),
      legend.text  = element_text(size = 6),
      legend.key.height = unit(0.3, "cm"),
      legend.key.width  = unit(0.15, "cm"),
      legend.position   = "bottom",
      plot.margin       = margin(0, 0, 0, 0)
    ) +
    guides(fill = guide_colorbar(barwidth = 25, barheight = 1)) +
    labs(title = title)
}





```


***Cross-Sector Analysis***
```{r}

# Below is a dataframe for the GARCH residuals for each sector, during the stable period
all_sector_garch_stable <- rbind(garch_communication_stable, garch_discretionary_stable, garch_staples_stable)

# Below is a dataframe for the GARCH residuals for each sector, during the pandemic period
all_sector_garch_pandemic <- rbind(garch_communication_pandemic, garch_discretionary_pandemic, garch_staples_pandemic)

all_sector_wide_stable <- get_wide_residuals(all_sector_garch_stable)
all_sector_wide_pandemic <- get_wide_residuals(all_sector_garch_pandemic)


all_tickers <- c(
  "META", "GOOGL", "NFLX", 
  "DIS", "WBD", "PARA", 
  "CMCSA", "EA", "T", 
  "VZ", "TMUS", "OMC",
  "AMZN", "TSLA","HD",
  "MCD", "LOW", "NKE",
  "SBUX", "MAR", "EBAY",
  "BBY", "ETSY", "MGM",
  "COST", "WMT", "KO",
  "PEP", "TGT", "KR",
  "TSN", "CLX", "STZ",
  "MNST", "SYY", "PG")


all_spearman_stable <-plot_spearman_correlation_ggplot_all_sector(all_sector_wide_stable, "Spearman Correlation Across All Sectors (Stable Period)", all_tickers)
all_spearman_pan <-plot_spearman_correlation_ggplot_all_sector(all_sector_wide_pandemic, "Spearman Correlation Across All Sectors (Pandemic Period)", all_tickers)



all_mi_stable <-plot_mi_mat_all_sector(all_sector_wide_stable, "Mutual Information Across All Sectors (Stable Period)", all_tickers)
all_mi_pan <-plot_mi_mat_all_sector(all_sector_wide_pandemic, "Mutual Information Across All Sectors (Pandemic Period)", all_tickers)

# For the future, the plot mutual information is calling the reshape function that filters for the lower half of the triangle



all_tdc_stable <-plot_tdc_mat_all_sector(all_sector_wide_stable, "TDC Across All Sectors (Stable Period)",tickers = all_tickers, percentile = 0.05)
all_tdc_pan <-plot_tdc_mat_all_sector(all_sector_wide_pandemic, "TDC Across All Sectors (Pandemic Period)",tickers = all_tickers, percentile = 0.05)






```


***Combining all cross sector plots***
```{r}



# Outputting the all of the plots for the communication services sector
all_spearman_stable <- plot_spearman_correlation_ggplot_all_sector(
  all_sector_wide_stable,
  "Stable Period",
  all_tickers
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(Spearman_stable_comm, position = NULL)
print(legend)

all_spearman_pan <- all_spearman_pan + theme(legend.position = "none")

all_spearman_pan <- plot_spearman_correlation_ggplot_all_sector(
  all_sector_wide_pandemic,
  "Pandemic Period",  
  all_tickers
) + theme(legend.position = "none")


comm_row <- plot_grid(
  all_spearman_stable, all_spearman_pan, nrow = 1
)




main_title <- ggdraw() + 
  draw_label("Spearman Correlation All Sectors ", fontface = "bold", size = 30, x = 0.35, hjust = 0)

combined_plot <-plot_grid(
  main_title,
  comm_row, my_legend,
  nrow = 3,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)
save_plot("all_sector_mi_plot.png", 
          combined_plot, 
          base_width = 25.2,   # in inches
          base_height = 14.8,  # in inches
          dpi = 300)


#################################################

library(patchwork)


# remove legend from the pandemic plot
p1 <- all_spearman_stable
p2 <- all_spearman_pan + theme(legend.position = "none")

combined_plot <-
  (p1 + p2) +
    plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation in All Sectors",
    theme = theme(
      plot.title        = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.title.position = "plot"
    )
  ) &
  theme(legend.position = "bottom")

ggsave(
  filename = "all_sector_spearman_plot2.png",
  plot     = combined_plot,
  width    = 25.2,
  height   = 14.8,
  dpi      = 300
)


#############################################




# Outputting the all of the plots for the communication services sector
all_mi_stable <- plot_mi_mat_all_sector(
  all_sector_wide_stable,
  "Stable Period",
  all_tickers
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(all_mi_stable, position = NULL)
print(legend)

all_mi_pan <- all_mi_pan + theme(legend.position = "none")

all_mi_pan <- plot_mi_mat_all_sector(
  all_sector_wide_pandemic,
  "Pandemic Period",  
  all_tickers
) + theme(legend.position = "none")


comm_row <- plot_grid(
  all_mi_stable, all_mi_pan, nrow = 1
)




main_title <- ggdraw() + 
  draw_label("Mutual Information in All Sectors ", fontface = "bold", size = 30, x = 0.35, hjust = 0)

combined_plot <-plot_grid(
  main_title,
  comm_row, my_legend,
  nrow = 3,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)


save_plot("all_sector_mi_plot.png", 
          combined_plot, 
          base_width = 25.2,   # in inches
          base_height = 14.8,  # in inches
          dpi = 300)

#################################################

library(patchwork)

# remove legend from the pandemic plot
p1 <- all_mi_stable
p2 <- all_mi_pan + theme(legend.position = "none")

combined_plot <-
  (p1 + p2) +
    plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Mutual Information in All Sectors",
    theme = theme(
      plot.title        = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.title.position = "plot"
    )
  ) &
  theme(legend.position = "bottom")

ggsave(
  filename = "all_sector_mi_plot.png",
  plot     = combined_plot,
  width    = 25.2,
  height   = 14.8,
  dpi      = 300
)






```



```{r}

# Outputting the all of the plots for the communication services sector
all_tdc_stable <- plot_tdc_mat_all_sector(
  all_sector_wide_stable,
  "Stable Period",
  all_tickers
)

# Extract a common legend from one plot (assuming all share the same legend)
my_legend <- ggpubr::get_legend(all_tdc_stable, position = NULL)
print(legend)

all_tdc_pan <- all_tdc_pan + theme(legend.position = "none")

all_tdc_pan <- plot_tdc_mat_all_sector(
  all_sector_wide_pandemic,
  "Pandemic Period",  
  all_tickers
) + theme(legend.position = "none")


comm_row <- plot_grid(
  all_tdc_stable, all_tdc_pan, nrow = 1
)




main_title <- ggdraw() + 
  draw_label("TDC in All Sectors ", fontface = "bold", size = 30, x = 0.35, hjust = 0)

combined_plot <-plot_grid(
  main_title,
  comm_row, my_legend,
  nrow = 3,
  rel_heights = c(0.05, 1, 1, 1, 0.1)
)


save_plot("all_sector_tdc_plot.png", 
          combined_plot, 
          base_width = 25.2,   # in inches
          base_height = 14.8,  # in inches
          dpi = 300)

#################################################

library(patchwork)

# remove legend from the pandemic plot
p1 <- all_tdc_stable
p2 <- all_tdc_pan + theme(legend.position = "none")

combined_plot <-
  (p1 + p2) +
    plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "TDC in All Sectors",
    theme = theme(
      plot.title        = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.title.position = "plot"
    )
  ) &
  theme(legend.position = "bottom")

ggsave(
  filename = "all_sector_tdc_plot.png",
  plot     = combined_plot,
  width    = 25.2,
  height   = 14.8,
  dpi      = 300
)

```

